---
title: "final-1-1"
author: "Kai McNamee"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(estimatr)
library(haven)
library(stargazer)
library(lubridate)
library(sjlabelled)
library(latticeExtra)
```

Use nationscape data with state -- is it a state level condition thing?

NOTES: add a column for 

```{r read-data message=FALSE}
### Nationscape data
# Collect the file names and select which ones we want

file_names_1 <- list.files("data/Nationscape-DataRelease_WeeklyMaterials_DTA/phase_1_v20200814/") %>% .[1:24]
file_names_2 <- list.files("data/Nationscape-DataRelease_WeeklyMaterials_DTA/phase_2_v20200814/") %>% .[1:26]

# Reading in all phase 1 and 2 weeks

phase_1 <- map_dfr(.x = file_names_1, 
                   ~read_dta(file = str_c("data/Nationscape-DataRelease_WeeklyMaterials_DTA/phase_1_v20200814/", ., "/", ., ".dta")) %>% 
                     remove_all_labels()) %>%   
  mutate(across(.cols = everything(), ~na_if(., 999))) %>% 
  mutate(across(.cols = everything(), ~na_if(., 888)))

phase_2 <- map_dfr(.x = file_names_2, 
                   ~read_dta(file = str_c("data/Nationscape-DataRelease_WeeklyMaterials_DTA/phase_2_v20200814/", ., "/", ., ".dta")) %>% 
                     remove_all_labels()) %>%   
  mutate(across(.cols = everything(), ~na_if(., 999))) %>% 
  mutate(across(.cols = everything(), ~na_if(., 888)))

# Join phase 1 and 2

ns_joined <- full_join(phase_1, phase_2)

# Select variables

ns_vars <- c(party = "pid3", 
             gender = "gender",
             race = "race_ethnicity",
             state = "state",
             fox = "news_sources_fox",
             cnn = "news_sources_cnn",
             asians = "group_favorability_asians")

ns_selected <- ns_joined %>% 
  select(response_id, start_date, ns_vars) %>% 
  mutate(party = case_when(party == 1 ~ "Democrat",
                           party == 2 ~ "Republican",
                           party == 3 ~ "Independent",
                           T ~ NA_character_),
         start_date = as_date(start_date)) %>% 
  rename(date = start_date)

### Covid data
# Read covid data from
# https://data.cdc.gov/Case-Surveillance/United-States-COVID-19-Cases-and-Deaths-by-State-o/9mfq-cb36

covid_states <- read_csv("data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv", 
                         col_types = cols(.default = col_guess())) %>% 
  rename(date = submission_date) %>%
  mutate(date = as_date(mdy(date))) %>% 
  filter(date < as_date("2020-08-01)")) %>% 
  select(date, state, tot_cases, new_case, tot_death, new_death)

### Lockdown data
# Read lockdown data from https://www.nytimes.com/interactive/2020/us/coronavirus-stay-at-home-order.html

lockdowns <- read_csv("data/nyt_stay_at_home.csv", col_types = cols(.default = col_guess())) %>% 
  select(Abbreviation, Date) %>% 
  rename(start_date = Date, state = Abbreviation) %>% 
  mutate(start_date = as_date(mdy(start_date)))
```

```{r dual-axis}
# Compare trends in attitudes towards asians and covid cases

cutoff <- as_date("2020-07-01")

ns_month <- ns_selected %>% 
  mutate(month_year = as_date(my(paste0(month(date), "-", year(date))))) %>% 
  select(month_year, asians) %>% 
  drop_na() %>% 
  group_by(month_year) %>%
  summarize(asians = mean(asians), .groups = "drop") %>% 
  filter(month_year < cutoff)
  # ggplot(aes(x = month_year, y = asians)) +
  #   geom_point()

covid_month <- covid_states %>% 
  mutate(month_year = as_date(my(paste0(month(date), "-", year(date))))) %>% 
  select(month_year, tot_cases) %>% 
  group_by(month_year) %>% 
  summarize(tot_cases = sum(tot_cases), .groups = "drop") %>% 
  filter(month_year < cutoff)
  # ggplot(aes(x = month_year, y = tot_cases)) +
  #   geom_point()

obj1 <- xyplot(asians ~ month_year, data = ns_month, typ = "l", lwd = 2, col="red")
obj2 <- xyplot(tot_cases ~ month_year, data = covid_month, typ = "l", lwd = 2, col="blue")
doubleYScale(obj1, obj2, add.ylab2 = TRUE, use.style=FALSE )


ns_month
covid_month
```

Notes
NS data is only collected on specific days that the survey is administered -- covid data is collected on different days. Does that mean I have to model covid data to get the number of cases for every day of the NS survey? (how to fill in the gaps??)

(This is actually wrong -- covid data exists for every day that survey data is collected)

```{r clean}

# NOTE: join is happening for group_by(date) totals --> do this join again such
# that each observation has US cases and state cases for respondents' state!!!

joined_data <- ns_selected %>% 
  left_join(covid_states %>% 
              group_by(date) %>% 
              summarize(tot_cases = sum(tot_cases),
                        new_case = sum(new_case),
                        tot_death = sum(tot_death),
                        new_death = sum(new_death),
                        .groups = "drop"), 
            by = "date") %>% 
  # mutate(tot_cases = ifelse(is.na(tot_cases) == TRUE, 0, tot_cases)) %>% 
  
  # Join lockdown data. 
  
  left_join(lockdowns, by = "state") %>% 
    
  # For case_when: 
  # 1. code lockdown as NA if the observation is before the pandemic started (ie when tot_cases == NA)
  # 2. code lockdown as FALSE if the respondent's state never implemented a
  #     lockdown (ie when tot_cases =/= NA, but lockdown start_date == NA)
  # 3. code lockdown as FALSE if the response is before the respondent's state
  #     implmented a lockdown (ie when tot_cases =/= NA, and lockdown start_date =/=
  #     NA)
  # 4. otherwise, code lockdown as TRUE (ie when tot_cases =/= NA, lockdown
  #     start_date =/= NA, and response date is after lockdown start_date)
  
  mutate(lockdown = case_when(is.na(tot_cases) == TRUE ~ NA, 
                              is.na(start_date) == TRUE ~ FALSE, 
                              date < start_date ~ FALSE,
                              T ~ TRUE)) %>% 
  drop_na(asians)

covid_date <- covid_states %>%
  group_by(date) %>%
  summarize(tot_cases = sum(tot_cases), .groups = "drop")
```


```{r initial-viz}
# ns_selected %>% 
#   drop_na() %>% 
#   group_by(date) %>% 
#   summarize(asians = mean(asians), .groups = "drop") %>% 
#   left_join(covid_states %>% 
#               group_by(date) %>% 
#               summarize(tot_cases = sum(tot_cases), .groups = "drop"), 
#             by = "date") %>% 
#   filter(date >= as_date("2020-01-22"))
  

# group_favorability_asians plot

joined_data %>% 
  drop_na(asians) %>% 
  group_by(date) %>% 
  summarize(asians = mean(asians),
            tot_cases = mean(tot_cases), .groups = "drop") %>%
  ggplot(aes(x = date, y = asians)) +
    geom_point(color = "#f8766d") + 
    geom_smooth(color = "#f8766d", method = "loess", formula = y ~ x) +
    theme_minimal() +
    labs(title = "Favorability towards Asians over time",
         x = "", y = "Favorability towards Asians") +
    scale_x_date(date_labels = "%B-%Y")

# tot_cases plot

joined_data %>% 
  drop_na(tot_cases) %>% 
  group_by(date) %>% 
  summarize(asians = mean(asians),
            tot_cases = mean(tot_cases), .groups = "drop") %>%
  ggplot(aes(x = date, y = tot_cases)) +
    geom_point(color = "#00bfc4") + 
    theme_minimal() +
    labs(title = "Total COVID cases over time",
         x = "", y = "Total COVID cases") +
    scale_x_date(date_labels = "%b-%Y")

labels <- c("asians" = "Favorability towards Asians", 
            "tot_cases" = "Total cases")

joined_data %>% 
  drop_na(asians, tot_cases) %>% 
  group_by(date) %>% 
  summarize(asians = mean(asians),
            tot_cases = mean(tot_cases), .groups = "drop") %>% 
  pivot_longer(cols = c(asians, tot_cases), names_to = "series", values_to = "values") %>% 
  ggplot(aes(x = date, y = values, color = series)) +
    geom_point(size = 0.5) +
    facet_wrap("series", scales = "free_y", labeller = as_labeller(labels)) +
    scale_x_date(date_labels = "%b-%Y") +
    scale_color_discrete(labels = as_labeller(labels), name = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    labs(title = "", x = "", y = "") 
```

```{r}
m1 <- lm(asians ~ date + tot_cases + new_case + tot_death + new_death + state + lockdown, joined_data %>% drop_na(tot_cases))
summary(m2)

m2 <- lm(asians ~ date + tot_cases + tot_death + lockdown, joined_data %>% drop_na(tot_cases))
m3 <- lm(asians ~ date + new_case + new_death + lockdown, joined_data %>% drop_na(tot_cases))
stargazer(m2, m3, type = "text")

```



